---
title: "Swirl Courses"
author: "ElisaRMA"
date: "3/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

According to the swirl website "swirl is a software package for the R programming language that turns the R console into an interactive learning environment. Users receive immediate feedback as they are guided through self-paced lessons in data science and R programming."

More information can be found on [https://swirlstats.com](https://swirlstats.com) 

## Install swirl

Swirl can be installed using the `install.packages` function
```{r}

install.packages("swirl")

```

To install any course within swirl, load the package and then install the course

```{r}
library(swirl)
install_from_swirl("course")
```

## Swirl Courses - Data Science Specialization - Coursera:

From the 5 courses within the Data Science Specialization - Foundations using R, 3 courses had its respective course on swirl. The codes used to install each one are available below. 

1. R Programming
```{r}

install_from_swirl("R Programming")

```

2. Getting and Cleaning Data
```{r}

install_from_swirl("Getting and Cleaning Data")

```

3. Exploratory Data Analysis
```{r}
 
install_from_swirl("Exploratory Data Analysis")

```

4. Statistical Inference
```{r}

install_from_swirl("Statistical Inference")

```

 
